---
title: "Simulation and Inferential Data Analysis"
output: html_document
---

##PART 1
 
This is the project for the statistical inference class. In it, you will use simulation to explore inference and do some simple inferential data analysis. The project consists of two parts:

1. A simulation exercise.
2. Basic inferential data analysis.

You will create a report to answer the questions. Given the nature of the series, ideally you'll use knitr to create the reports and convert to a pdf. (I will post a very simple introduction to knitr). However, feel free to use whatever software that you would like to create your pdf. 

Each pdf report should be no more than 3 pages with 3 pages of supporting appendix material if needed (code, figures, etcetera).



In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.


Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponential(0.2)s.  You should

1. Show the sample mean and compare it to the theoretical mean of the distribution.

```{r}
setwd("~/Desktop/Coursera/StatInf/courseProject")
library(lattice)
library(gplots)

lambda <- 0.2
n <- 40
simulations <- 1:2000
set.seed(55) #this is needed to make the exercise reproducible (can be any number)
means <- data.frame(x = sapply(simulations, function(x) {mean(rexp(n, lambda))}))
```
```{r}
mean(means$x) # sampling mean = 5.015019
(1/lambda) # theoretical mean = 5
```


The distribution is centered at 5.015019 (sampling mean) which is similar to the theoretical mean which is 5

2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

```{r}
var(means$x) # sampling var = 0.6212932
sd(means$x) # sampling sd = 0.7882215
((1/lambda)/sqrt(n))^2 # theoretical var = 0.625
(1/lambda)/sqrt(n) # theoretical sd = 0.7905694
```

Sampling variance and standard deviation are 0.6212932 and 0.7882215 respectively (sampling var and sd) 

Theoretical variance and standard deviation are 0.625 and 0.7905694 respectively (theoretical var and sd)

3. Show that the distribution is approximately normal.

A plot is useless, some plot don't look normal but they are normal and some look normal but they aren't normal, instead of a plot I'll use a normality test. After performing a test a plot can give some support to the test.

```{r}
shapiro.test(means$x) # p-value = 2.972e-08
```

p < 0.05 means the data is represented by a normal curve, which is the case

```{r, echo=FALSE}
tmp <- hist(means$x, freq=TRUE, main="Data histogram and normal curve",
  	col="lightgreen", axes = FALSE, xlab='Mean', yaxt='n', ylab='Frequency')
axis(2, pos = 2.5)
axis(1,at=seq(0,8,1), pos=0)

multiplier <- tmp$counts / tmp$density
mydensity <- density(means$x)
mydensity$y <- mydensity$y * multiplier[1]

myx <- seq(2.5, 8, length.out= 100)
mymean <- mean(means$x)
mysd <- sd(means$x)

normal <- dnorm(x = myx, mean = mymean, sd = mysd)
lines(myx, normal * multiplier[1], col = "blue", lwd = 3)
```

##PART 2

Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package.

1. Load the ToothGrowth data and perform basic exploratory data analysis 

```{r}
data(ToothGrowth)
nrow(ToothGrowth)
ncol(ToothGrowth)
str(ToothGrowth)
```

```{r}
ToothGrowth$dose
ToothGrowth$dose = factor(ToothGrowth$dose, levels=c(0.5,1.0,2.0),
                          labels=c("low","med","high"))
attach(ToothGrowth)
table(supp,dose)
```

```{r, echo=FALSE}
bwplot(len ~ dose | supp)
```



2. Provide a basic summary of the data.

```{r}
summary(ToothGrowth)
aggregate(len,list(supp,dose), mean)
aggregate(len,list(supp,dose), sd)
```

3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose. (Use the techniques from class even if there's other approaches worth considering)

Working with a factor (dose), hence Anova will be useful.

```{r}                 
anova <- aov(len ~ supp * dose, data=ToothGrowth)
TukeyHSD(anova)
```

```{r, echo=FALSE}
plotmeans(len~interaction(supp,dose,sep=" "),connect=list(c(1,3,5),c(2,4,6)),col=c("red","darkgreen"),xlab="Dose and Supp Combination",main = "Interaction plot with 95% confidence intervals")
```

4. State your conclusions and the assumptions needed for your conclusions. 

Assumptions: The sample components are independent and are identically distributed.

Conclussion: Tukey contrast reveals there is no statistical evidence for other 'noisy' variables which are relevant, therefore changes in dose and size (independent variables) won't affect length (dependent variable)
